---
title: "Transform clean data for analysis"
author: "Daniel Cerritos"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
library(here)
```

## Load data

```{r load data}
(files <- list.files(path = here("output", "clean_data"), full.names = TRUE))

multilocation_2017 <- read_csv(files[4])
urbana_2018 <- read_csv(files[5])
iowa <- read_csv(files[3])
```

## Eliminate observations

```{r wrong observations}
# eliminate data because they had weigth errors
multilocation_2017[c(12, 20), c(11, 12)] <- NA
iowa[1, 18] <- NA

# no other 1a variety for comparison and we did not used in 2018 either 
# we decided to remove it 
multilocation_2017 <- multilocation_2017 %>% 
  filter(!variety == "LD12-8677")

# eliminate row 2 from variety LD10-10219 in Monmouth because the stand counts in those rows are strange
# there are to high, probably error when writing
multilocation_2017[which(multilocation_2017$variety == "LD10-10219" & 
                           multilocation_2017$location == "Monmouth"), 
                   c("stand.VC_row2", "stand.V2_row2") ] <- NA

```

## Transform multilocation 2017 data
- Stand count from multilocation field trials in 2017 was measured in 1 m segments in the two central rows
- I would transform stand count into plants/m2 by averaging the number of plants in the two rows and then divide by the area
- row width was 0.76 m and segment measured was 1 m for 2017 data, area = 0.76 m2  
- yield (bu/acre) would be calculated based on the moisture and weight (lb) 
- formula comes from one of the combines 
- then multiply by 62.77 to transform it to kg/ha


```{r transform  2017 data}
multilocation_2017 <- multilocation_2017 %>% 
  rowwise() %>% 
  mutate(vc_plants.m2 = mean(c(vc_row1, vc_row2), na.rm = TRUE),
         v2_plants.m2 = mean(c(v2_row1, v2_row2), na.rm = TRUE)
         ) %>% 
  mutate(vc_plants.m2 = round((vc_plants.m2/0.762)), 
         v2_plants.m2 = round((v2_plants.m2/0.762))
         )

# yield for Monmouth and Urbana, they both had 5.1 m (17 ft) long rows
mon.urb.yield <- multilocation_2017 %>%
  filter(!location == "Orr") %>% 
  mutate(yield_kg.ha = 
           round((43560*weight_lb*(100-moisture)/
                    (5*17)/(60*87))*62.77, digits = 1)
         )
# Orr rows were 6.9 m (22 ft) long
orr.yield <- multilocation_2017 %>%
  filter(location == "Orr") %>% 
  mutate(yield_kg.ha = 
           round((43560*weight_lb*(100-moisture)/
                    (5*22)/(60*87))*62.77, digits = 1)
         )
multilocation_2017 <- rbind(mon.urb.yield, orr.yield)
```

## Transform Urbana and Iowa 2018 data
- Urbana and Boone 2018 stand count was measured in 3.048 m segments (10 ft were flagged). 
- row width was 0.76 m and segment measured was 3.048 m for 2018 
- area = 2.31648
- results from combine for 2018 already had yield (bu/acre) so I multiplied by 62.77 to transform to kg/ha

```{r transform 2018 data}
ft.to.m <- function(x) (round(x/2.323))

urbana_2018 <- urbana_2018 %>% 
  rowwise() %>% 
  mutate(vc_plants.m2 = mean(c(vc_row1, vc_row2)), 
         v2_plants.m2 = mean(c(v2_row1, v2_row2)),
         v4_plants.m2 = mean(c(v4_row1, v4_row2)),
         r8_plants.m2 = mean(c(r8_row1, r8_row2))
         ) %>%
  mutate_at(c("vc_plants.m2", 
              "v2_plants.m2", 
              "v4_plants.m2", 
              "r8_plants.m2"), 
            ft.to.m
            )

iowa <- iowa %>% 
  rowwise() %>% 
  mutate(vc_plants.m2 = mean(c(vc_row1, vc_row2)), 
         v1_plants.m2 = mean(c(v1_row1, v1_row2)),
         r2_plants.m2 = mean(c(r2_row1, r2_row2)),
         r8_plants.m2 = mean(c(r8_row1, r8_row2))
         ) %>%
  mutate_at(c("vc_plants.m2", 
              "v1_plants.m2", 
              "r2_plants.m2", 
              "r8_plants.m2"), 
            ft.to.m
            )

kg.ha <- function(x) (round(x*62.77))

urbana_2018 <- urbana_2018 %>% 
  mutate(yield.kg_ha = kg.ha(yield_bu.acre))

iowa <- iowa %>% 
  mutate(yield.kg_ha = kg.ha(yield_bu.acre))
```


## Create whole-plot column 
- experimental design was a split-plot so whole plot error (block * variety) is included in the model 
- block is nested in location

```{r whole plot}
multilocation_2017 <- multilocation_2017 %>% 
    mutate(whole_plot = paste(location, block, variety, sep = ":")
           )
whole.plot <- function(data){
  data %>% 
    mutate(whole_plot = paste(block, variety, sep = ":")
           )
}
urbana_2018 <- whole.plot(urbana_2018)
iowa <- whole.plot(iowa)
```

## Save data

```{r save data}

write_csv(multilocation_2017, path = here("output", 
                                          "transform_data", 
                                          "2019-09-27_multilocation-2017_transform-data.csv"))

write_csv(urbana_2018, path = here("output", 
                                   "transform_data", 
                                   "2019-09-27_urbana-2018_transform-data.csv"))

write_csv(iowa, path = here("output", 
                            "transform_data", 
                            "2019-09-27_iowa-2018_transform-data.csv"))

```


