---
title: "Clean raw data"
author: "Daniel Cerritos"
date: "9/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Cleaning data for analysis of efficacy field trials of ethaboxam seed treatment  
in combination with resistant varieties. Data from three locations were collected 
in 2017 field trials and recorded in sheets in Field Book 2017. In 2018,  
just one location was used and data recorded in the tablet.   
We also received data from a 2018 field trial in Iowa. Stand count and yield data was collected.
Varieties used in Illinois were previously screen for Phytphthora sojae resistance.

```{r library}
library(tidyverse)
library(readxl)
library(here)
library(naniar)
```

## Load data

```{r list all files}
# file names were not changed, same name as found in box 
(raw.files <- list.files(here("data"), full.names = TRUE)) 
```

```{r load data, echo=FALSE}
# Resistance screening 
tray.test <- raw.files[14] %>% 
  excel_sheets() %>%  # divided in multiple sets (sheets)
  set_names() %>% 
  map(read_excel, path = raw.files[14]) 

layer.test <- raw.files[8] %>% 
  excel_sheets() %>%
  set_names() %>% 
  map(read_excel, path = raw.files[8])

# stand counts are in different excel sheets, yields in different files 
# Monmouth 
mon.stand.vc <- read_excel(raw.files[11], sheet = 1)
mon.stand.v2 <- read_excel(raw.files[11], sheet = 2)
mon.yield <- read_excel(raw.files[3])

# Orr
orr.stand.vc <- read_excel(raw.files[12], sheet = 1)
orr.stand.v2 <- read_excel(raw.files[12], sheet = 2)
orr.yield <- read_excel(raw.files[1])

# Urbana 
urb.stand.vc <- read_excel(raw.files[13], sheet = 1)
urb.stand.v2 <- read_excel(raw.files[13], sheet = 2)
urb.stand.r8 <- read_excel(raw.files[13], sheet = 4) # there was final stand
urb.yield <- read_csv(raw.files[10])

# Urbana 2018
urb.18.stand <- read_csv(raw.files[2])
urb.18.yield <- read_excel(raw.files[4], skip = 2)

# Iowa 2018
iowa <- read_excel(raw.files[5])
```

## P. sojae screening

```{r clean screening data}

str(layer.test) # only need two sheets

layer.test$`Group 2`$`#plants deat...5`


layer.test.1 <- layer.test$Group1 %>% 
  select(line = Line, 
         rep = Rep, 
         scale_dorrance = `Scale dorrance`, 
         seedlings_alive = `#plants alive`, 
         seedlings_dead = `#plants deat`, 
         )

layer.test.2 <- layer.test$`Group 2` %>% 
  select(line = Line...1, 
         rep = Rep...2, 
         scale_dorrance = `Scale dorrance...3`,
         seedlings_alive =`#plants alive...4`, 
         seedlings_dead = `#plants deat...5`
         )

layer.test <- bind_rows(layer.test.1, layer.test.2)

write_csv(layer.test, path = here("output", 
                                  "clean_data", 
                                  "2019-10-15_psojae-screening_layer-test_clean-data.csv")
          )


tray.test <- bind_rows(tray.test) %>% 
  rename(id = `PRR#`, 
         line = Strain, 
         seedling_number = Plant, 
         lesion_length.cm = Lenglesion)

write_csv(tray.test, path = here("output", 
                                 "clean_data", 
                                 "2019-10-15_psojae-screening_tray-test_clean-data.csv")
          )
```

## Multilocation field trials 2017 

```{r clean monmouth}
str(mon.stand.vc)

mon.row1 <- mon.stand.vc %>% 
  filter(row_number() %% 2 == 1) %>% # select even rows
  dplyr::select(vc_row1 = stand)

mon.row2 <- mon.stand.vc %>% 
  filter(row_number() %% 2 == 0) %>% 
  dplyr::select(vc_row2 = stand)
  
mon.stand.vc <- cbind(mon.row1, mon.row2)

str(mon.stand.v2) # row stands in different column, also includes a disease rating

mon.stand.v2 <- mon.stand.v2 %>% 
  select(v2_row1 = `Stand Row 1`,
         v2_row2 = `Stand Row 2`, 
         damping_off = `Phytophthora (1-5)`)


str(mon.yield)

mon.yield <- mon.yield %>% 
  select(sub_plot = plot, 
         weight_lb = Weight...12, 
         moisture = Moisture...13)

view(mon.yield)

# add an extra zero to first eight sub_plots to arrange them in order
rows.yield <- mon.yield[1:8, ] 

rows.yield$sub_plot <- str_replace_all(rows.yield$sub_plot,
                                       c("17SM001" = "17SM0001",
                                         "17SM002" = "17SM0002",
                                         "17SM003" = "17SM0003",
                                         "17SM004" = "17SM0004",
                                         "17SM005" = "17SM0005",
                                         "17SM006" = "17SM0006",
                                         "17SM007" = "17SM0007",
                                         "17SM008" = "17SM0008")
                                        )
mon.yield <- mon.yield[-c(1:8), ] # eliminate old rows

mon.yield <- mon.yield %>% 
  rbind(rows.yield) %>% 
  arrange(sub_plot) 

mon.var.trt.data <- read_excel(raw.files[11], sheet = 4) %>% 
  select(sub_plot = plot_name, 
         block = rep, 
         variety = var, 
         treatment = treat) # add rest variety and treatmetn info 



monmouth <- bind_cols(mon.var.trt.data, mon.stand, mon.yield)

monmouth <- monmouth %>% 
  select(-sub_plot1) %>% 
  mutate(location = "Monmouth")
```

```{r clean orr}
str(orr.stand.vc)

orr.row1 <- orr.stand.vc %>% 
  filter(row_number() %% 2 == 1) %>% 
  select(sub_plot = `plot_ name`, 
         block, 
         variety, 
         treatment, 
         vc_row1 = Stand1)

orr.row2 <- orr.stand.vc %>% 
  filter(row_number() %% 2 == 0) %>% 
  select(vc_row2 = Stand1)

orr.stand.vc <- bind_cols(orr.row1, orr.row2)

orr.stand.v2 <- orr.stand.v2 %>%
  select(v2_row1 = row1, 
         v2_row2 = Row2, 
         damping_off = `Phytoptora death`)

orr.yield <- orr.yield %>% 
  select(weight_lb = Weight, 
         moisture = Moisture)

orr.info <- read_excel(raw.files[12], sheet = 4) %>% 
  select(sub_plot = row_id)


orr <- bind_cols(orr.info, 
                 orr.stand.vc, 
                 orr.stand.v2, 
                 orr.yield) %>% 
  select(-sub_plot1) %>% 
  mutate(location = "Orr")
```

```{r clean urbana}
str(urb.stand.vc)
str(urb.stand.v2)
str(urb.yield)

urb.stand.vc <- urb.stand.vc %>% 
  select(plot_name, row, stand) %>% 
  spread(row, stand) %>% 
  rename(vc_row1 = `1`, 
         vc_row2 = `2`)

urb.stand.v2 <- urb.stand.v2 %>% 
  select(v2_row1 = `Stand Row 1`, 
         v2_row2 = `Stand Row 2`, 
         damping_off = `Phytophthora (1-5)`)

urb.stand.r8 <- urb.stand.r8 %>% 
  select(r8_row1 = `Stand Row 1`, 
         r8_row2 = `Stand Row 2`)


# Yield file for Urbana 2017 did not have all the plot names
# added manually in Excel using haverst sequence and the field map 

urb.yield <- urb.yield %>% 
  select(sub_plot = Plot, 
         weight_lb = Weight, 
         moisture = Moisture, 
         ) %>% 
  filter(!sub_plot == "Fill") %>% 
  arrange(sub_plot)

# add variety and treatment info
urb.info <- read_excel(raw.files[13], sheet = 5) %>% 
  select(sub_plot = plot_name, 
         block = rep, 
         variety = var,
         treatment = treat)


urbana <- bind_cols(urb.info, 
                    urb.stand.vc, 
                    urb.stand.v2, 
                    urb.stand.r8, 
                    urb.yield
                    )

urbana <- urbana %>% 
  select(-c(plot_name, sub_plot1)) %>% 
  mutate(location = "Urbana")
```

```{r combine 2017 data}

multilocation_2017 <- bind_rows(urbana, monmouth, orr)

# change "_" for "-" in variety name to be consistant with naming
multilocation_2017$variety <- str_replace_all(multilocation_2017$variety, 
                                     c("LD07_3395bf" = "LD07-3395bf", 
                                       "LD10_10219" = "LD10-10219",
                                       "LD11_10069" = "LD11-10069" , 
                                       "LD11_13802R2" = "LD11-13802R2", 
                                       "LD11_7311" = "LD11-7311", 
                                       "LD12_15156R1a" = "LD12-15156R1a", 
                                       "LD12_8677" = "LD12-8677", 
                                       "LD13_14071R2" = "LD13-14071R2")
                                     ) 
write_csv(multilocation_2017, 
          path = here("output", 
                      "clean_data", 
                      "2019-9-25_multilocation-2017_clean-data.csv")
          )
```

## Urbana and Iowa field trials 2018

```{r clean urbana 2018}

str(urb.18.stand)
view(urb.18.stand)

urb.18.stand <- urb.18.stand %>% 
  filter(!is.na(Block) # eliminate fill plots
         ) %>% 
  arrange(`Plot ID`) %>% 
  mutate(row = case_when( 
  Row %in% c(82, 78, 74, 70, 66, 62, 58, 54, 50, 46, 42, 38) ~ "row1", 
  TRUE ~ "row2" # change row numbers to 1 and 2                
  )) %>%
  select(`Source Stock`, 
         block = Block, 
         vc = Stand1, 
         v2 = Stand2, 
         v4 = Stand3, 
         r8 = `Final Stand`, 
         row) %>% 
  group_by(`Source Stock`, block) %>% 
  pivot_wider(names_from = row, 
              values_from = c(vc, v2, v4, r8)
              ) %>% 
  separate(`Source Stock`, 
           sep = "_", 
           into = c("variety", "treatment")
           ) 

view(urb.18.yield)

urb.18.yield <- urb.18.yield[-c(73:81), ] # remove fill rows

urb.18.yield <- urb.18.yield %>% 
  select(weight = Weight, 
         moisture = Moisture, 
         yield_bu.acre = Yield)

urbana_2018 <- bind_cols(urb.18.stand, urb.18.yield)

write_csv(urbana_2018, 
          path = here("output", 
                      "clean_data", 
                      "2019-9-25_urbana-2018_clean-data.csv")
          )
```

```{r clean iowa}
# Iowa data was sent organized

names(iowa) <- c("location", "variety", "treatment", 
                 "variety_treatment","block", "sub_plot", 
                 "vc_row1", "vc_row2", 
                 "v1_row1", "v1_row2",
                 "r2_row1", "r2_row2", 
                 "r8_row1", "r8_row2",
                 "damping_off", "moisture",
                 "weight_kg", "yield_bu.acre")

iowa$treatment <- str_replace_all(iowa$treatment,
                                  c("Y" = "intego",
                                    "N" = "control"
                                    )
                                  )

write_csv(iowa, 
          path = here("output", 
                      "clean_data", 
                      "2019-9-25_iowa-2018_clean-data.csv")
          )
```





