---
title: "Seed treatment field trials analysis"
author: "Daniel Cerritos"
date: "12/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      dev = "png",
                      cache = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

```

```{r library, message=FALSE}
library(tidyverse) 
library(here)
library(lme4) 
library(lmerTest) 
library(emmeans) 
library(multcompView) 
library(broom) 
library(cowplot) 
library(ggResidpanel)
library(ggsignif)
theme_set(theme_bw())
```

# Import data files  


```{r import field trials data}
seed.treatments.data <- list.files(path = here("output", "transform_data"), 
                         full.names = TRUE)
  
illinois.2017 <- read.csv(seed.treatments.data[[2]]) 
urbana.2018 <- read.csv(seed.treatments.data[[3]])
iowa <- read.csv(seed.treatments.data[[1]])
```



# Multilocation field trials 2017 


Fit mixed model for 2017 field trials. The experimental design was a split-plot so the whole-plot error is  
included in model. 

```{r model vc 2017}
illinois.2017$block <- as.factor(illinois.2017$block) # set block as factor

fm.illinois.vc <- lmer(vc_plants.m2 ~ location*variety * treatment + (1|block:location) + 
                         (1|whole_plot), data = illinois.2017)
anova(fm.illinois.vc) 
```

```{r vc means, fig.height=2, fig.width=6}
emmeans(fm.illinois.vc, ~ treatment) # increase of 12.9%
emmeans(fm.illinois.vc, ~ treatment) %>% 
  plot(comparisons = TRUE) 
```

Dots represent means and bars are the confidence intervals. Arrows are for comparisons, arrows that overlap are not significant different. 


```{r model v2 2017}
fm.illinois.v2 <- lmer(v2_plants.m2 ~ location*variety*treatment + (1|block:location) + 
                         (1|whole_plot), data = illinois.2017)
anova(fm.illinois.v2)
```

```{r v2 means 2017, fig.height=2, fig.width=6}
emmeans(fm.illinois.v2, ~ treatment)# increase of 8.2%
emmeans(fm.illinois.v2, ~ treatment) %>%   
  plot(comparisons = TRUE) 
```

```{r model yield 2017}
fm.illinois.yield <- lmer(yield_kg.ha ~ location * variety * treatment + (1|block:location) +
                            (1|whole_plot), data = illinois.2017)
anova(fm.illinois.yield)
```

```{r yield mean 2017, fig.height=2, fig.width=6}
emmeans(fm.illinois.yield, ~ treatment) # 0.6 %
emmeans(fm.illinois.yield, ~ treatment) %>% 
  plot(comparisons = TRUE)
```

The seed treatment significantly increased early stands (VC and V2), but no effect was  observed for yield. 



## Analysis by location 

Location was significant for all variables measured. Location by variety was significant for stand at VC and yield.   
An analysis by location would be perform to see if specific varieties perform better in a particular location.

```{r 2017 data by location}
# change whole_plot column by removing location nested in block
whole.plot <- function(data){
  data %>% 
    mutate(whole_plot = paste(block, variety, sep = ":"))
}

monmouth <- illinois.2017 %>% 
  filter(location == "Monmouth") %>% 
  whole.plot() 

orr <- illinois.2017 %>% 
  filter(location == "Orr") %>% 
  whole.plot()

urbana <- illinois.2017 %>% 
  filter(location == "Urbana") %>% 
  whole.plot()
```



## Monmouth 


```{r fit models monmouth}
fm.mon.vc <- lmer(vc_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = monmouth)
fm.mon.v2 <- lmer(v2_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = monmouth)
fm.mon.yield <- lmer(yield_kg.ha ~ variety*treatment + (1|block) + (1|whole_plot), data = monmouth)

anova(fm.mon.vc) 
anova(fm.mon.v2)
anova(fm.mon.yield) 
```

Treatment effect on stand (VC and V2), but not on yield. Variety is significant for VC and yield.


### Treatment comparisons 

I will focus in VC because both treatment and variety were significant. 

```{r vc monmouth}
emmeans(fm.mon.vc, ~ treatment|variety) %>% 
  cld.emmGrid()
```

LD11-10069 (Rps1c, moderate) is the only variety with significant differences between control and treatment.


### Yield comparisons

No effect of seed treatment on yield so I would only compare yield between varieties.

```{r yield monmouth}
emmeans(fm.mon.yield, ~ variety) %>% 
  cld.emmGrid()
```

Varieties with Rps don't seem to perform better than varieties with no gene (rps). The variety with low levels of paritial resistance (LD12-15156R1a) has the lowest yield.  


### Contrasts

I will group varieties by gene and by levels of partial resistance. Then do some contrasts betweem the seed treatment and untreated control to see the overrall effect of the seed treatment by type of resistance.

```{r means monmouth}
m.mon.vc <- emmeans(fm.mon.vc, ~ variety:treatment) 
m.mon.yield <- emmeans(fm.mon.yield, ~ variety:treatment)
```

```{r custom contrast}
custom_contrasts <- function(lsmeans){
  # by Rps
  cons.1 <- list(
    "t.rps - c.rps" = c(-0.5, -0.5, rep(0, 5), 0.5, 0.5, rep(0, 5)),
    "t.Rps1c - c.Rps1c" = c(0, 0, -1/3, 0, 0, -1/3, -1/3, 0, 0, 1/3, 0, 0, 1/3, 1/3),
    "t.Rps1k - c.Rps1k" = c(rep(0, 3), -0.5, -0.5, rep(0, 5), 0.5, 0.5, 0, 0))
  # by partial resistance level
  cons.2 <- list(
    "t.High - c.High" = c(-0.5, rep(0, 5), -0.5, 0.5, rep(0, 5), 0.5), 
    "t.Moderate - c.Moderate" = c(0, rep(-0.25, 4), rep(0, 3), rep(0.25, 4), 0, 0), 
    "t.Low - c.Low" = c(rep(0, 5), -1, rep(0, 6), 1, 0)
  )
  cus.con1 <- contrast(lsmeans, cons.1, adjust = "mvt") # adjustment for mutiple testing
  cus.con2 <- contrast(lsmeans, cons.2, adjust = "mvt") 
  con.both <- c(cus.con1, cus.con2)
  confi_con <- map(con.both, summary, infer = TRUE) # confidence intervals
  df_1 <- as.data.frame(confi_con[[1]])
  df_2 <- as.data.frame(confi_con[[2]])
  df_all <- rbind(df_1, df_2)
  df_all
}
```


I will keep focusing only in stand VC and yield.

```{r effect VC monmouth}
custom_contrasts(m.mon.vc) %>% 
  knitr::kable()
```

Significant differences for Rps1c.


```{r effect yield monmouth}
custom_contrasts(m.mon.yield) %>% 
    knitr::kable()
```

No significant differences found for yield. The control for varieties with high levels of partial resistance had
higher yield.


I would do another contrast comparing yields between levels of partial resistance (PR). Rps didn't seem to perform 
better than rps, but high PR seem to perform better.

```{r pr level monmouth}
pr_contrast <- function(lsmean){
  cons.3 <- list(
  "High - Low" = c(0.5, 0, 0, 0, 0, -1, 0.5),
  "High - Moderate" = c(0.5, -0.25, -0.25, -0.25, -0.25, 0, 0.5),
  "Moderate - Low" = c(0, 0.25, 0.25, 0.25, 0.25, -1, 0)
  )
  cus.con.3 <- contrast(lsmean, cons.3, adjust = "mvt")
  conf_con <- summary(cus.con.3, infer = TRUE)
  df_con <- as.data.frame(conf_con)
  df_con
}

emmeans(fm.mon.yield, ~ variety) %>% 
  pr_contrast() %>% 
  knitr::kable()
```

High level PR have higher yield than low, but no difference with moderate. No difference between moderate and low.



## Orr

```{r fit models orr}
fm.orr.vc <- lmer(vc_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = orr)
fm.orr.v2 <- lmer(v2_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = orr)
fm.orr.yield <- lmer(yield_kg.ha ~ variety*treatment + (1|block) + (1|whole_plot), data = orr)
anova(fm.orr.vc) 
anova(fm.orr.v2) 
anova(fm.orr.yield) 
```


### Treatment comparison  


```{r vc orr}
emmeans(fm.orr.vc, ~ treatment|variety) %>% 
  cld.emmGrid()
```

Only one variety (LD12-15156R1a) had a significant difference between seed treatment and control.


### Yield comparison


```{r yield orr}
emmeans(fm.orr.yield, ~ variety) %>% 
  cld.emmGrid() 
```

Same as Monmouth, variety with no resistance gene has the higher yield.


### Contrasts


```{r effect vc orr}
emmeans(fm.orr.vc, ~ variety:treatment) %>% 
  custom_contrasts() %>% 
  knitr::kable()
```

Significant differences for Rps1k. Moderate and low levels also have a significant differences.


```{r}
emmeans(fm.orr.yield, ~ variety:treatment) %>% 
  custom_contrasts() %>% 
  knitr::kable()
```

No significant differences for yield, but again the untreated control has higher yield for varieties with high levels of partial resistance.


```{r pr level orr}
emmeans(fm.orr.yield, ~ variety) %>% 
  pr_contrast() %>% 
  knitr::kable()
```

High partial reistance varieties have higher yields than both moderate and low.



## Urbana


```{r fit models urbana}
fm.urb.vc <- lmer(vc_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana)
fm.urb.v2 <- lmer(v2_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana)
fm.urb.r8 <- lmer(r8_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana)
fm.urb.yield <- lmer(yield_kg.ha ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana)

anova(fm.urb.vc)
anova(fm.urb.v2)
anova(fm.urb.r8) 
anova(fm.urb.yield) 
```

Treatment is significant for the final stand at R8.



A significant interaction was observed for VC in Urbana.

```{r interaction urbana, fig.width=8, fig.height=4}
emmip(fm.urb.vc, treatment ~ variety)
```

Seems that seed treatment is having an effect in both rps and Rps1k. 


### Treatment comparison 


```{r vc urbana}
emmeans(fm.urb.vc, ~ treatment|variety) %>% 
  cld.emmGrid()
```

No significant difference for two of the the three Rps1c varieties. 


### Yield comparison

```{r yield urbana}
emmeans(fm.urb.yield, ~ variety) %>% 
  cld.emmGrid()
```

A Rps1c variety has the highest yield.


### Contrasts


```{r effect vc urbana}
emmeans(fm.urb.vc, ~ variety:treatment) %>% 
  custom_contrasts %>% 
  knitr::kable()
```

Significatn effect for rps and Rps1k. Moderate and low also had a significant difference.e


```{r effect yield urbana}
emmeans(fm.urb.yield, ~ variety:treatment) %>% 
  custom_contrasts() %>% 
  knitr::kable()
```

No difference for yield, but the untreated control for Rps1c and high levels of partial had higher yields.


```{r pr level urbana}
emmeans(fm.urb.yield, ~ variety) %>% 
  pr_contrast() %>% 
  knitr::kable()
```

High and moderate partial resistance have higher yields than low levels. No difference between high and moderate.




## Figure 1


```{r add resistance info, echo=FALSE}
# add resistance information (Rps and partial resistance levels) for figures
resistance_info <- function(dataf) {
  rps <- c("LD07-3395bf","LD10-10219","LD13-13478R1a")
  Rps1k <- c("LD11-13802R2", "LD11-7311", "H2512NX", "H2862NX")
  low <- c("LD12-15064R1a", "LD12-15129R1a", "LD13-13478R1a", "LD12-15156R1a")
  high <- c("LD07-3395bf", "LD13-14071R2", "C2888RX", "C3140RX")
  
  dataf %>% 
    mutate(
      rps = case_when(
        variety %in% c("LD07-3395bf","LD10-10219","LD13-13478R1a") ~ "rps",
        variety %in% Rps1k ~ "Rps1k",
        TRUE ~ "Rps1c"),
      pr_level = case_when(
        variety %in% low ~ "Low",
        variety %in% high  ~ "High",
        TRUE ~ "Moderate"))
}

illinois.2017 <- resistance_info(illinois.2017)
```

```{r fig 1 labels, echo=FALSE}
# treatment comparisons 
lab.fig.1 <- data.frame(variety = c("LD11-10069", "LD12-15156R1a",
                                "LD07-3395bf", "LD10-10219", 
                                "LD12-15156R1a", "LD11-13802R2",
                                "LD11-7311"), 
                    rps= c("Rps1c", "Rps1c", "rps", "rps", "Rps1c", 
                           "Rps1k", "Rps1k"),
                    location = c("Monmouth", "Orr", "Urbana", "Urbana", 
                                 "Urbana", "Urbana", "Urbana"),
                    label = c("*", "*", "*", "*", "*", "*", "*"), 
                    vc_plants.m2 = c(28, 29, 35, 36, 34, 34, 32), 
                    treatment = c("intego", "intego", "intego", "intego",
                                  "intego", "intego","intego")) 
# contrasts
con.fig.1 <- data.frame(location = c("Monmouth", "Orr", 
                                     "Urbana", "Urbana"), 
                        rps = c("Rps1c", "Rps1k", 
                                "rps", "Rps1k"), 
                        start = c("LD11-10069", "LD11-13802R2", 
                                  "LD07-3395bf", "LD11-13802R2"), 
                        end = c("LD13-14071R2", "LD11-7311", 
                                "LD10-10219", "LD11-7311"), 
                          y = c(34, 34, 41, 39), 
                        label = c("*", "*", "*", "*"), 
                        treatment = c("intego", "intego", 
                                      "intego", "intego"))
# shorter varieties names
short.names <- c("LD10-10219" = "LD10-10", 
                 "LD13-13478R1a" = "LD13-13",
                 "LD07-3395bf" = "LD07-33", 
                 "LD12-15064R1a" = "LD12-150",
                 "LD12-15129R1a" = "LD12-151a",
                 "LD11-10069" = "LD11-10",
                 "LD13-14071R2" = "LD13-14",
                 "LD12-15156R1a" = "LD12-151", 
                 "LD11-13802R2" = "LD11-13",
                 "LD11-7311" = "LD11-73") 

```

```{r fig 1, warning = FALSE, fig.width = 7, fig.height = 6, echo=FALSE}
figure1 <- illinois.2017 %>% 
  ggplot(aes(x = variety, y = vc_plants.m2)) +
  geom_point(aes(color = treatment), 
               position = position_jitterdodge(dodge.width = 0.5), 
               alpha = 0.2) +
    stat_summary(fun.data = "mean_cl_boot", 
                 position = position_dodge(width = 0.5),
                 aes(color = treatment), alpha = 0.7) + 
  facet_grid(location ~ rps, scales = "free", space = "free_x") +
  labs(y = expression('Stand (plants m'^-2*')')) +
  theme_cowplot() +
  panel_border() +
  theme(panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80"), 
        legend.position = "bottom") +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("UTC", "ST"), 
                     name = element_blank()) +
  scale_x_discrete("Variety", labels = short.names) +
  geom_text(data = lab.fig.1, aes(label=label), 
            size = 5, color = "black") +
  geom_signif(data = con.fig.1, 
              aes(xmin = start, xmax = end, annotations=label, y_position = y),
              textsize = 5, vjust = -0.2, 
              manual = TRUE) +
  scale_y_continuous(limits = c(15, 45))

ggsave("cerritos-garcia1.tiff", plot = figure1, dpi = 300, path = here("docs", "figure"))
```

## Figure 2

Box plots of the distribution of yield of varieties with different levels of partial resistance in multilocation field
experiment in Illinois 2017. The center lines show the medians and box limits indicate the 25th and 75th percentiles
and whiskers extend 1.5 times the interquartile range (IQR). Outliers are represented by black solid dots


```{r fig 2, warning = FALSE, fig.width = 6.5, fig.height = 5.5, echo = FALSE}
illinois.2017$pr_level <- factor(illinois.2017$pr_level, 
                                 levels = c("High", 
                                            "Moderate", 
                                            "Low"))
figure2 <- illinois.2017 %>% 
  ggplot(aes(x = fct_reorder(variety, -yield_kg.ha), 
             y = yield_kg.ha, fill= pr_level)) +
  geom_jitter(alpha= 0.2, width = 0.4) +
  geom_boxplot(alpha = 0.5) +
  facet_grid(location ~ ., scales = "free") +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73"), 
                    name = element_blank()) +
  theme_cowplot() +
  panel_border() +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80")) +
  labs(y = expression('Yield (kg ha'^-1*')'),
         x = "Variety") +
  scale_x_discrete("Variety", labels = short.names) +
  stat_summary(fun.y="mean", geom="point", size=1.2,
                 position=position_dodge(width=0.75), color="black")

ggsave("cerritos-garcia2.tiff", plot = figure2, dpi = 300, path = here("docs", "figure"))
```



# Urbana and Boone field trials in 2018


## Urbana, IL

```{r fit models urbana 2018}
urbana.2018$block <- as.factor(urbana.2018$block)

fm.urb18.vc <- lmer(vc_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana.2018)
fm.urb18.v2 <- lmer(v2_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana.2018)
fm.urb18.v4 <- lmer(v4_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana.2018)
fm.urb18.r8 <- lmer(r8_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana.2018)
fm.urb18.yield <- lmer(yield_kg.ha ~ variety*treatment + (1|block) + (1|whole_plot), data = urbana.2018)

anova(fm.urb18.vc)
anova(fm.urb18.v2) 
anova(fm.urb18.v4)  
anova(fm.urb18.r8) 
anova(fm.urb18.yield)
```

Treatment effect significant in the four stand counts and also for yield. Variety is only significant for stand at VC
and yield.


### Treatment comparison 

```{r treatment urbana 18, fig.height=2, fig.width=6}
# overrall treatment effect
emmeans(fm.urb18.vc, ~ treatment) %>% #17.82%
  plot(comparison = TRUE)
emmeans(fm.urb18.v2, ~ treatment) %>% #18.61
  plot(comparison = TRUE)
emmeans(fm.urb18.v4, ~ treatment) %>% #18.35
    plot(comparison = TRUE)
emmeans(fm.urb18.r8, ~ treatment) %>% #15.6
    plot(comparison = TRUE)
emmeans(fm.urb18.yield,  ~ treatment) %>% #7.84
    plot(comparison = TRUE)
```


I will focus in VC and yield like in 2017 analysis. 

```{r vc urbana 18}
# seed treatment effect within variety
emmeans(fm.urb18.vc, ~ treatment|variety) %>% 
  cld.emmGrid()
```

Varieties LD07-3395bf(rps, High) and LD11-7311 (1k, moderate) have significant differences.


```{r yield trt urb18}
emmeans(fm.urb18.yield, ~ treatment|variety) %>% 
  cld.emmGrid()
```

The two varieties with high moderate levels, LD07-3395bf(rps) and LD13-14071R2 (1c)  
have significant differences 


### Yield comparison

```{r yield var}
emmeans(fm.urb18.yield, ~ variety) %>% 
  cld.emmGrid()
```

High level partial resistance varieties have higher yields. A Rps1c variety had the highest yield.


### Constrasts 

Same contrasts between untreated control and seed treatment by each type of resistance (resistance gene or partial resistance level).


```{r contrasts urb18}
custom_contrast18 <- function(lsmeans){
  # by Rps
  cons.1.18 <- list(
  "t.rps - c.rps" = c(-1/3, -1/3, rep(0, 5), -1/3, 0, 1/3, 1/3, rep(0, 5), 1/3, 0),
  "t.Rps1c - c.Rps1c" = c(rep(0, 4), -0.25, -0.25, -0.25, 0, -0.25, rep(0, 4), 0.25, 0.25, 0.25, 0, 0.25),
  "t.Rps1k - c.Rps1k" = c(rep(0, 2), -0.5, -0.5, rep(0, 7), 0.5, 0.5, rep(0, 5))
)
  # by partial resistance level
  cons.2.18 <- list(
  "t.High - c.High" = c(-0.5, rep(0, 7), -0.5, 0.5, rep(0, 7), 0.5), 
  "t.Moderate - c.Moderate" = c(0, rep(-1/3, 3), rep(0, 6), rep(1/3, 3), rep(0, 5)), 
  "t.Low - c.Low" = c(rep(0, 4), rep(-1/4, 4), rep(0, 5), rep(1/4, 4), 0)
)
  cus.con1 <- contrast(lsmeans, cons.1.18, adjust = "mvt") 
  cus.con2 <- contrast(lsmeans, cons.2.18, adjust = "mvt") 
  con.both <- c(cus.con1, cus.con2)
  confi_con <- map(con.both, summary, infer = TRUE) 
  df_1 <- as.data.frame(confi_con[[1]])
  df_2 <- as.data.frame(confi_con[[2]])
  df_all <- rbind(df_1, df_2)
  df_all
}
```

```{r effect vc urb18}
emmeans(fm.urb18.vc, ~ variety:treatment) %>% 
  custom_contrast18() %>% 
  knitr::kable()
```

Significant differences for rps and Rps1k. Significant differences for moderate levels.


```{r effect yield urb18}
emmeans(fm.urb18.yield, ~ variety:treatment) %>% 
  custom_contrast18()%>% 
  knitr::kable()
```

Significant differences for rps and Rps1c. Significant differences for high levels of partial resistance.


```{r pr level urb18}
pr_contrast18 <- function(lsmean){
  cons.3.18 <- list(
  "High - Low" = c(0.5, 0, 0, 0, -1/3, -1/3, 0, -1/3, 0.5),
  "High - Moderate" = c(0.5, -0.25, -0.25, -0.25, 0, 0, -0.25, 0, 0.5),
  "Moderate - Low" = c(0, 0.25, 0.25, 0.25, -1/3, -1/3, 0.25, -1/3, 0)
)
  cus.con.3 <- contrast(lsmean, cons.3.18, adjust = "mvt")
  conf_con <- summary(cus.con.3, infer = TRUE)
  df_con <- as.data.frame(conf_con)
  df_con
}
```

```{r pr level yield urb18}
emmeans(fm.urb18.yield, ~ variety) %>% 
  pr_contrast18%>% 
  knitr::kable()
```

High level partial resistnace has higher yields than both moderate and low levels.



## Figure 3

```{r fig 3A, warning=FALSE, echo=FALSE}
urbana.2018 <- resistance_info(urbana.2018)
urbana.2018$pr_level <- factor(urbana.2018$pr_level, levels = c("High", 
                                                                "Moderate", 
                                                                "Low"))

lab.fig.3A <- data.frame(variety = c("LD07-3395bf", "LD11-7311"), 
                         rps= c("rps", "Rps1k"),
                         label = c("*", "*"), 
                         treatment = c("intego", "intego"), 
                         vc_plants.m2 = c(30, 30))

con.fig.3A <- data.frame(rps = c("rps", "Rps1k"), 
                        start = c("LD07-3395bf", "LD11-13802R2"), 
                        end = c("LD13-13478R1a", "LD11-7311"), 
                        y = c(35, 35), 
                        label = c("*", "*"), 
                        treatment = c("intego", "intego"))

fig.3A <- urbana.2018 %>% 
  ggplot(aes(x = variety, y = vc_plants.m2)) +
  geom_point(aes(color = treatment), 
             position = position_jitterdodge(dodge.width = 0.5), 
             alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", 
               position = position_dodge(width = 0.5),
               aes(color = treatment), alpha = 0.7) +
  facet_grid(. ~ rps, scales = "free", space = "free_x") +
  labs(y = expression('Stand (plants m'^-2*')')) +
  theme_cowplot() +
  panel_border() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(), 
        panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80")) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("NTC", "ST"), 
                     name = element_blank()) +
  geom_text(data = lab.fig.3A, aes(label=label), 
            size = 5, color = "black") +
  geom_signif(data = con.fig.3A, 
              aes(xmin = start, xmax = end, 
                  annotations=label, 
                  y_position = y),
              textsize = 5, vjust = -0.2, 
              manual = TRUE) +
  scale_y_continuous(limits = c(10, 40))
```

```{r fig 3B, echo=FALSE, warning=FALSE}

lab.fig.3B <- data.frame(variety = c("LD07-3395bf", "LD13-14071R2"), 
                         rps= c("rps", "Rps1c"),
                         label = c("*", "*"), 
                         yield_kg.ha = c(4300, 4500), 
                         treatment = c("intego", "intego"))

con.fig.3B <- data.frame(rps = c("rps", "Rps1c"), 
                        start = c("LD07-3395bf", "LD12-15064R1a"), 
                        end = c("LD13-13478R1a", "LD13-14071R2"), 
                        y = c(4800, 5000), 
                        label = c("*", "*"), 
                        treatment = c("intego", "intego"))

fig.3B <- urbana.2018  %>% 
  ggplot(aes(x = variety, y = yield_kg.ha)) +
  geom_point(aes(color = treatment), 
             position = position_jitterdodge(dodge.width = 0.5), 
             alpha = 0.2) +
  facet_grid(. ~ rps, scales = "free", space = "free_x") +
  stat_summary(fun.data = "mean_cl_boot", 
               position = position_dodge(width = 0.5),
               aes(color = treatment), alpha = 0.7) +
  labs(y = expression('Yield (kg ha'^-1*')')) +
  theme_cowplot() +
  panel_border() +
  theme(legend.position = "bottom",
    legend.box.margin=margin(-10,-10,-10,-10),
    axis.title.x = element_blank(),
    strip.text = element_blank(),
    axis.text.x = element_text(size = 8),
    panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80"), 
        ) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("UTC", "ST"), 
                     name = element_blank()) +
  geom_text(data = lab.fig.3B, aes(label=label), 
            size = 5, color = "black") +
  geom_signif(data = con.fig.3B, 
              aes(xmin = start, xmax = end, annotations=label, y_position = y),
              textsize = 5, vjust = -0.2, 
              manual = TRUE) +
  scale_x_discrete(labels = short.names)+
  scale_y_continuous(limits = c(2000, 5500))
```

```{r fig 3C, echo=FALSE}
fig.3C <- urbana.2018 %>% 
  ggplot(aes(x = fct_reorder(variety, -yield_kg.ha), 
             y = yield_kg.ha, fill= pr_level)) +
  geom_jitter(alpha= 0.2, width = 0.4) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73"), 
                    name = element_blank()) +
  theme_cowplot() +
  panel_border() +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80"), 
        axis.text.x = element_text(size = 10)) +
  labs(y = expression('Yield (kg ha'^-1*')'),
         x = "Variety") +
  scale_x_discrete("Variety", labels = short.names) +
  stat_summary(fun.y="mean", geom="point", size=1.2,
                 position=position_dodge(width=0.75), color="black")
```

```{r figure 3, echo=FALSE, fig.width = 7, fig.height = 7}
figure3 <- cowplot::plot_grid(fig.3A, fig.3B, fig.3C,
                   ncol = 1, 
                   align = "v", 
                   labels = "AUTO", 
                   axis = "lr", 
                   rel_heights = c(1, 1, 1.3))

figure3
ggsave("cerritos-garcia3.tiff", plot = figure3, dpi = 300, path = here("docs", "figure"))
```



## Boone, Iowa 

```{r boone fit models}
iowa$block <- as.factor(iowa$block)

fm.iowa.vc <- lmer(vc_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = iowa)
fm.iowa.v1<- lmer(v1_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = iowa)
fm.iowa.r2 <- lmer(r2_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = iowa)
fm.iowa.r8 <- lmer(r8_plants.m2 ~ variety*treatment + (1|block) + (1|whole_plot), data = iowa)
fm.iowa.yield <- lmer(yield_kg.ha ~ variety * treatment + (1|block) + (1|whole_plot), data = iowa)

anova(fm.iowa.vc) 
anova(fm.iowa.v1)
anova(fm.iowa.r2)  
anova(fm.iowa.r8) 
anova(fm.iowa.yield)

```

Treatment was significant for all stands and yield. Variety was not significant for 


### Treatment effect

```{r emmeans boone treatment, fig.height=2, fig.width=6}
emmeans(fm.iowa.vc, ~ treatment) %>%  #12.25
  plot(comparisons = TRUE)
emmeans(fm.iowa.v1,  ~ treatment) %>% #9.62
  plot(comparisons = TRUE)
emmeans(fm.iowa.r2,  ~ treatment) %>%  #11.10
    plot(comparisons = TRUE)
emmeans(fm.iowa.r8,  ~ treatment) %>% #12.01
    plot(comparisons = TRUE)
emmeans(fm.iowa.yield,  ~ treatment) %>%  #8.1
    plot(comparisons = TRUE)
```

All stand counts and yield are significanlty increased with the seed treatment.


### Yield comparisons


```{r yield iowa}
emmeans(fm.iowa.yield, ~ variety) %>% 
  cld.emmGrid()
```

No difference between varieties for yield.


### Contrasts

```{r iowa contrasts}
custom_contrastiw <- function(lsmeans){
  # compare effect of seed treatment by Rps
cons.1.iw <- list(
  "t.Rps1c - c.Rps1c" = c(-0.2, -0.2, -0.2, 0, 0, -0.2, -0.2, 0.2, 0.2, 0.2, 0, 0, 0.2, 0.2),
  "t.Rps1k - c.Rps1k" = c(0, 0, 0, -0.5, -0.5, 0, 0, 0, 0, 0, 0.5, 0.5, 0, 0)
)
  cons.2.iw <- list(
  "t.High - c.High" = c(0, 0, -0.5, 0, 0, 0, -0.5, 0, 0, 0.5, 0, 0, 0, 0.5), 
  "t.Moderate - c.Moderate" = c(-1/5, -1/5, 0, -1/5, -1/5, -1/5, 0, 1/5, 1/5, 0, 1/5, 1/5, 1/5, 0)
)
  
  cus.con1 <- contrast(lsmeans, cons.1.iw, adjust = "mvt") 
  cus.con2 <- contrast(lsmeans, cons.2.iw, adjust = "mvt") 
  con.both <- c(cus.con1, cus.con2)
  confi_con <- map(con.both, summary, infer = TRUE) 
  df_1 <- as.data.frame(confi_con[[1]])
  df_2 <- as.data.frame(confi_con[[2]])
  df_all <- rbind(df_1, df_2)
  df_all
}
```

```{r iowa means}
emmeans(fm.iowa.vc, ~ variety:treatment) %>% 
  custom_contrastiw() %>% 
  knitr::kable()
```

Difference for Rps1c and high levels of partial resistance

```{r effect vc}
emmeans(fm.iowa.yield, ~ variety:treatment) %>% 
  custom_contrastiw() %>% 
  knitr::kable()
```

Difference for Rps1c and high levels of partial resistance

```{r pr levels iowa}
pr_contrastiw <- function(lsmean){
 cons.3.iw <- list(
  "High - Moderate" = c(-1/5, -1/5, 0.5, -1/5, -1/5, -1/5, 0.5)
)
  cus.con.3 <- contrast(lsmean, cons.3.iw, adjust = "mvt")
  conf_con <- summary(cus.con.3, infer = TRUE)
  df_con <- as.data.frame(conf_con)
  df_con
}
```

```{r con pr iowa}
emmeans(fm.iowa.yield, ~ variety) %>% 
  pr_contrastiw() %>% 
  knitr::kable()
```

No difference between high and moderate.


## Figure 4



```{r fig 4A, echo=FALSE}
iowa <- resistance_info(iowa)
lab.fig.4A <- data.frame(rps = "Rps1c", 
                         variety = "C3140RX", 
                         vc_plants.m2 = 13.5, 
                         label = "*", 
                         treatment = "intego")
con.fig.4A <- data.frame(rps = "Rps1c", 
                         start = "AG28x7",
                         end = "NK3195X",  
                         y = 15.5, 
                         label = "*", 
                         treatment = "intego")

fig.4A <- iowa %>% 
  ggplot(aes(x = variety, y = vc_plants.m2)) +
  geom_point(aes(color = treatment), 
             position = position_jitterdodge(dodge.width = 0.5), 
             alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", 
               position = position_dodge(width = 0.5),
               aes(color = treatment), alpha = 0.7) +
  facet_grid(. ~ rps, scales = "free", space = "free_x") +
  labs(y = expression('Stand (plants m'^-2*')')) +
  theme_cowplot() +
  panel_border() +
  theme(axis.text.x = element_text(size = 10),
        panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80"), 
        axis.title.x = element_blank()) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("UTC", "ST"), 
                     name = element_blank()) +
  geom_text(data = lab.fig.4A, aes(label=label), 
            size = 5, color = "black")+
   geom_signif(data = con.fig.4A, 
              aes(xmin = start, xmax = end, annotations=label, y_position = y),
              textsize = 5, vjust = -0.2, 
              manual = TRUE)+ 
  scale_y_continuous(limits = c(8, 18))
```

```{r fig 4B, echo=FALSE}
iowa <- iowa %>% 
  mutate(pr_level = forcats::fct_inorder(pr_level))

fig.4b <- iowa %>% 
  ggplot(aes(x = fct_reorder(variety, -yield_kg.ha), 
             y = yield_kg.ha, fill= pr_level)) +
  geom_jitter(alpha= 0.2, width = 0.4) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73"), 
                    name = element_blank()) +
  theme_cowplot() +
  panel_border() +
  theme(
        panel.grid.major.y = element_line(linetype = "dotted", 
                                          color = "grey80"), 
        axis.text.x = element_text(size = 10)) +
  labs(y = expression('Yield (kg ha'^-1*')'),
         x = "Variety") +
  stat_summary(fun.y="mean", geom="point", size=1.2,
                 position=position_dodge(width=0.75), color="black")
```

```{r fig 4, fig.width = 7, fig.height = 4.4, echo=FALSE}
figure4 <- cowplot::plot_grid(fig.4A, fig.4b,
                   ncol = 1, 
                   align = "v", 
                   labels = "AUTO", 
                   axis = "lr" )

figure4
ggsave("cerritos-garcia4.tiff", plot = figure4,  width=7, height=4, dpi = 300, path = here("docs", "figure"))
```




# Model assumptions


Diagnostic plots to check on model assumptions 


```{r assumptions models 2017}
# 2017
models.2017 <- list(fm.mon.vc, fm.mon.v2, fm.mon.yield, 
                    fm.orr.vc, fm.orr.v2, fm.orr.yield,
                    fm.urb.vc, fm.urb.v2, fm.urb.yield)
map(models.2017, resid_panel) #residuals normally distributed and homogenicity of variance
```

```{r assumptions mmodels 2018}
# 2018
models.2018 <- list(fm.urb18.vc, fm.urb18.v2, fm.urb18.v4, 
                    fm.urb18.r8, fm.urb18.yield, 
                    fm.iowa.vc, fm.iowa.v1, fm.iowa.r2, 
                    fm.iowa.r8, fm.iowa.yield)
map(models.2018, resid_panel) #residuals normally distributed and homogenicity of variance
```



